Cover Letter

### 2017-12-22 paper/sections/3_posix_overheads.tex

% Effects on performance
Figure~\ref{fig:overhead-a} shows the effect of journaling with different
\oldcomment{segment sizes; the larger the segment size the bigger that the
writes into the object store are}\newcomment{dispatch sizes, normalized to 1
client that creates 100K files with journaling off (about 654 creates/sec). In
this case a dispatch size of 10 degrades performance more because the metadata
server is overloaded with requests and cannot spare cycles to manage concurrent
segments.  Although the ``no journal" curve appears flat, it is actually a
slowdown of about \(0.3\times\) per concurrent client; this slowdown is a
result of the peak throughput of a single metadata server, which we found to be
about 3000 operations per second}.  The trade-off for better performance is
memory consumption because a larger \oldcomment{segment}\newcomment{dispatch}
size uses more space for buffering.  When journaling is on, the metadata server
periodically stops serving requests to flush ({\it i.e.} apply journal updates)
to the metadata store.  The journal overhead is sufficient enough to slow down
metadata throughput but not so much as to overwhelm the bandwidth of the object
store.

>>>>

% Effects on performance
Figure~\ref{fig:overhead-a} shows the effect of journaling with different
\oldcomment{segment sizes; the larger the segment size the bigger that the
writes into the object store are}\newcomment{dispatch sizes, normalized to 1
client that creates 100K files with journaling off (about 654 creates/sec). In
this case a dispatch size of 10 degrades performance the most because the
metadata server is overloaded with requests and cannot spare cycles to manage
concurrent segments.  Tuning and parameter sweeps show that a dispatch size of
10 is the worst and that larger sizes approach a dispatch size of 1; for all
future journal experiments we use a dispatch size of 40 which is a more
realistic configuration.  Although the ``no journal" curve appears flat, it is
actually a slowdown of about \(0.3\times\) per concurrent client; this slowdown
is a result of the peak throughput of a single metadata server, which we found
to be about 3000 operations per second}.  The trade-off for better performance
is memory consumption because a larger
\oldcomment{segment}\newcomment{dispatch} size uses more space for buffering.
When journaling is on, the metadata server periodically stops serving requests
to flush ({\it i.e.} apply journal updates) to the metadata store.  The journal
overhead is sufficient enough to slow down metadata throughput but not so much
as to overwhelm the bandwidth of the object store.

### 2017-12-22 paper/sections/6_results.tex

{\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
mechanisms that rely on poorly scaling data structures have large slowdowns for
the larger number of creates. For example, RPCs has a \(90\times\) slowdown
because it relies on the internal CephFS directory data structures. It is a
well-known problem that directory data structures do not scale when creating
files in the same directory~\cite{ren:sc2014-indexfs} and any mechanism that
uses these data structures will experience similar slowdowns. Other mechanisms
that write events to a journal) experience a much less drastic slowdown because
the journal data structure does not need to be scanned for every operation.
Events are written to the end of the journal without even checking the validity
({\it e.g.}, if the file already exists for a create), which is another form of
relaxed consistency because the file system assumes the application has
resolved conflicting updates in a different way.

@@

\textbf{Takeaway}: measuring the mechanisms individually shows that their
overheads and costs can differ {\it by orders of magnitude}. We also show that
some mechanisms, like Nonvolatile Apply, are not worthwhile as currently
implemented.

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
      \newcomment{parallel creates on clients}
      }\label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
      \newcomment{block interference}
      }\label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
      \newcomment{syncing to global namespace}
      }\label{fig:slowdown-sync}
  \end{subfigure}
\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
RPCs; \texttt{create} is the throughput of clients creating files in-parallel
and writing updates locally; \texttt{create+merge} includes the time to merge
updates at the metadata server. }\oldcomment{shows the cost of merging client
journals at the metadata server. Shipping and merging journals of
updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
are less messages and consistency/durability code paths are bypassed. (b) shows
how the allow/block API isolates directories from interfering clients. (c) is
the slowdown of a single client syncing updates to the global namespace. The
inflection point is the trade-off of frequent updates vs. larger journal
files.}
\label{fig:use-cases}
\end{figure*}

First we show how we can build application-specific subtrees by composing
mechanisms and that this approach results in more a scalable global namespace.
We start with clients creating files in private directories because this
workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
But the workload also appears in cloud workloads~\cite{xiao:socc15-shardfs},
where systems like Hadoop use the file abstraction to exchange work units to
workers or to indicate when jobs
complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
services a flash crowd of creates across all directories as shown in
Figure~\ref{fig:overhead-creates}.  The workload is clients creating 100K files
in private directories in the same global namespace.

@@

\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.\\

To show the contention at the metadata server, in Figure~\ref{fig:mergescale}
we scale the number of clients for the RPC per operation strategy and the
decoupled namespace strategy. For RPCs, we increase the number of clients
making requests to the metadata server and for decoupled (the merge bars), we
increase the number of concurrent journal merge requests. So for the 30 clients
data point, we are measuring the operations per second for 30 client journals
that land on the metadata server at the same time. The ``create \& merge" bar
adds the time it takes for clients to generate the journal of events in
parallel. The experiment is run three times but the standard deviation is too
small to see.

@@

\textbf{Takeaway}: the ability to couple well-established techniques to
specific applications allows the global namespace to scale further and perform
better. In our case, RPCs overwhelm the system but decoupling/shipping a
journal of updates improves scalability by 2\(\times\).

@@

Next we show how Cudele can be programmed to block interfering clients using
the same problematic workload from Figure~\ref{fig:overhead-b}. In this
workload, clients create files in private directories and a separate client
interferes by creating more files in each directory. This introduces false
sharing and the metadata server revokes capabilities on directories touched by
the interfering client. While HPC tries to avoid these situations with
workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
happens in distributed file systems when users unintentionally access
directories in a shared file system.  We only scale to 15 clients because the
performance variability of a nearly overloaded metadata server, in our case 18
clients, is too high.

Figure~\ref{fig:block-allow} plots the overhead of the slowest client and the
error bars are the standard deviation of 3 runs.  ``Interfere (Fig 3c)" is the
result from Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API
to return \texttt{-EBUSY} to interfering clients; and ``Isolated" is the
baseline performance without an interfering. Each subtree has RPCs and
Stream enabled.  Results are normalized to the runtime of a single
client creating files in a directory. Note that IndexFS has a similar behavior
to ``interfere" with leases, except clients block.

We draw three conclusions: (1) clients that use the API to block interfering
clients  get the same performance as isolated clients, (2) there is a
negligible effect on performance for the extra work the metadata server does to
return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.

\textbf{Takeaway}: the API lets users isolate directories when applications
need better and more reliable performance. Blocking updates is an effective way of
controlling consistency.

@@

The final use case shows how the API gives users fine-grained control of the
consistency semantics.  The use case is that users often leverage the file
system to check the progress of jobs using \texttt{ls} even though this
operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
eshel:fast10-panache} . The number of files or size of the files is indicative
of the progress. This practice is not too different from systems that use the
file system to manage the progress of jobs; Hadoop writes to temporary files,
renames them when complete, and creates a ``DONE" file to indicate to the
runtime that the task did not fail and should not be re-scheduled on another
node. In this scenario, Cudele users will not see the progress of decoupled
namespaces since their updates are not globally visible.  To help users
judge the progress of their jobs, Cudele clients have a ``namespace sync" that
sends batches of updates back to the global namespace at regular intervals.

% Does this implementation need to be described in the implementation?
Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
client writing updates to a decoupled namespace and pausing to send updates to
the metadata server. The error bars are the standard deviations of 3 runs. We
scale the namespace sync interval to show the trade-off of frequently pausing
or writing large logs of updates.  We use an idle core to log the updates and
to do the network transfer. The client only pauses to fork off a background
process, which is expensive as the address space needs to copied. The
alternative is to pause the client completely and write the update to disk but
since this implementation is limited by the speed of the disk, we choose the
memory-to-memory copy of the fork approach.

\textbf{Takeaway}: syncing namespace updates has up to a 9\% overhead and can
be tuned depending on the users preference but, more importantly, the API gives
users fine-grain control of their consistency/durability to support current
practices or experimental workflows.

>>>>

{\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
mechanisms that rely on poorly scaling data structures \oldcomment{have large
slowdowns}\newcomment{can slowdown performance by {\it orders of magnitude}}
for the larger number of creates. For example, RPCs has a \(90\times\) slowdown
because it relies on the internal CephFS directory data structures. It is a
well-known problem that directory data structures do not scale when creating
files in the same directory~\cite{ren:sc2014-indexfs} and any mechanism that
uses these data structures will experience similar slowdowns. Other mechanisms
that write events to a journal) experience a much less drastic slowdown because
the journal data structure does not need to be scanned for every operation.
Events are written to the end of the journal without even checking the validity
({\it e.g.}, if the file already exists for a create), which is another form of
relaxed consistency because the file system assumes the application has
resolved conflicting updates in a different way.

@@

\oldcomment{\textbf{Takeaway}: measuring the mechanisms individually shows that their
overheads and costs can differ {\it by orders of magnitude}. We also show that
some mechanisms, like Nonvolatile Apply, are not worthwhile as currently
implemented.}

\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
      \newcomment{parallel creates on clients}
      }\label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
      \newcomment{block interference}
      }\label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
      \newcomment{syncing to global namespace}
      }\label{fig:slowdown-sync}
  \end{subfigure}
\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
RPCs; \texttt{create} is the throughput of clients creating files in-parallel
and writing updates locally; \texttt{create+merge} includes the time to merge
updates at the metadata server. }\oldcomment{shows the cost of merging client
journals at the metadata server. Shipping and merging journals of
updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
are less messages and consistency/durability code paths are bypassed. (b) shows
how the allow/block API isolates directories from interfering clients. (c) is
the slowdown of a single client syncing updates to the global namespace. The
inflection point is the trade-off of frequent updates vs. larger journal
files.}
\label{fig:use-cases}
\end{figure*}

@@

First we show how we can build application-specific subtrees by composing
mechanisms and that \oldcomment{this approach}\newcomment{coupling
well-established techniques to specific applications} results in a more a
scalable global namespace.  We start with clients creating files in private
directories because this workload is heavily studied in
HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
cloud workloads~\cite{xiao:socc15-shardfs}, where systems like Hadoop use the
file abstraction to exchange work units to workers or to indicate when jobs
complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
services a flash crowd of creates across all directories as shown in
Figure~\ref{fig:overhead-creates}.  The workload is clients creating 100K files
in private directories in the same global namespace.

@@

\oldcomment{\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.\\}

To show the contention at the metadata server, in Figure~\ref{fig:mergescale}
we scale the number of clients for the RPC per operation strategy and the
decoupled namespace strategy. For RPCs, we increase the number of clients
making requests to the metadata server and for decoupled we increase the number
of concurrent clients operating in parallel. So for the 30 clients data point,
we are measuring the operations per second for 30 client journals that land on
the metadata server at the same time. The ``create + merge" bar adds the time
it takes for clients to generate the journal of events in parallel. The
experiment is run three times but the standard deviation is too small to see.

@@

\oldcomment{\textbf{Takeaway}: the ability to couple well-established techniques to
specific applications allows the global namespace to scale further and perform
better. In our case, RPCs overwhelm the system but decoupling/shipping a
journal of updates improves scalability by 2\(\times\).}

@@

Next we show how Cudele can be programmed to block interfering
clients\oldcomment{ using the same problematic workload from
Figure~\ref{fig:overhead-b}. In this workload, c}\newcomment{, which lets
applications control isolation to get better and more reliable performance.
C}lients create files in private directories while another client interferes by
creating \oldcomment{more}\newcomment{1000} files in each directory. This
introduces false sharing and the metadata server revokes capabilities on
directories touched by the interfering client. While HPC tries to avoid these
situations with workflows~\cite{zheng:pdsw2014-batchfs,
zheng:pdsw2015-deltafs}, it still happens in distributed file systems when
users unintentionally access directories in a shared file system.
\oldcomment{We only scale to 15 clients because the performance variability of
a nearly overloaded metadata server, in our case 18 clients, is too high.}

\newcomment{We switch back to clients doing RPCs per operation with journaling
enabled, to mirror the setup from the problem presented in
Figure~\ref{fig:overhead-b}.} Figure~\ref{fig:block-allow} plots the overhead
of the slowest client \oldcomment{and the error bars are the standard deviation
of 3 runs.  ``Interfere (Fig 3c)" is the result from
Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API to return
\texttt{-EBUSY} to interfering clients; and ``Isolated" is the baseline
performance without an interfering.  Each subtree has RPCs and Stream enabled.
Results are normalized to the runtime of a single client creating files in a
directory.  Note that IndexFS has a similar behavior to ``interfere" with
leases, except clients block.  }\newcomment{normalized to 1 client that creates
100K files in isolation (about 513 creates/sec). ``Interference" and ``no
interference" is the performance with and without an interfering client
touching files in every directory, respectively. The goal is to explicitly
isolate clients so that performance is similar to the ``no interfere" curve,
which has lower slowdowns (on average, \(1.42\times\) per client compared to
\(1.67\times\) per client for ``interfere") and less variability (on average, a
standard deviation of 0.06 compared to 0.44 for ``interfere").  ``Block
interference" uses the Cudele API to block interfering clients with
\texttt{-EBUSY}. That curve shows slowdowns similar to ``interfere" for smaller
clusters because the overhead to reject requests is more evident when the
metadata server is underloaded; but the slowdown (\(1.34\times\) per client)
and variability (0.09) look very similar to ``no interfere" for larger
clusters. Clients can block interfering clients to get the same performance as
isolated clients but there is a non-negligible overhead for rejecting requests
when the metadata server is not operating at peak efficiency.}

\oldcomment{We draw three conclusions: (1) clients that use the API to block interfering
clients  get the same performance as isolated clients, (2) there is a
negligible effect on performance for the extra work the metadata server does to
return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.

\textbf{Takeaway}: the API lets users isolate directories when applications
need better and more reliable performance. Blocking updates is an effective way of
controlling consistency.}

@@

The final use case shows how the API gives users fine-grained control of the
consistency semantics\newcomment{ to support current practices and experimental
workflows}. \oldcomment{The use case is that u}\newcomment{U}sers often
leverage the file system to check the progress of jobs using \texttt{ls} even
though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
eshel:fast10-panache} . The number of files or size of the files is indicative
of the progress. This practice is not too different from systems that use the
file system to manage the progress of jobs; Hadoop writes to temporary files,
renames them when complete, and creates a ``DONE" file to indicate to the
runtime that the task did not fail and should not be re-scheduled on another
node. In this scenario, Cudele users will not see the progress of decoupled
namespaces since their updates are not globally visible.  To help users judge
the progress of their jobs, Cudele clients have a ``namespace sync" that sends
batches of updates back to the global namespace at regular intervals.

% Does this implementation need to be described in the implementation?
Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
client writing \newcomment{1 million} updates to a decoupled namespace and
pausing to send updates to the metadata server. The error bars are the standard
deviations of 3 runs. We scale the namespace sync interval to show the
trade-off of frequently pausing or writing large logs of updates.  We use an
idle core to log the updates and to do the network transfer. The client only
pauses to fork off a background process, which is expensive as the address
space needs to copied. The alternative is to pause the client completely and
write the update to disk but since this implementation is limited by the speed
of the disk, we choose the memory-to-memory copy of the fork approach.

As expected, syncing namespace updates too frequently has the highest overhead
(up to 9\% overhead if done every second). The optimal sync interval for
performance is 10 seconds because larger intervals must write more updates to
disk and network. For the 25 second interval, the client only pauses 3-4 times
but each sync writes about 278 thousand updates at once, which is a 678MB
journal.

\oldcomment{\textbf{Takeaway}: Syncing namespace updates has up to a 9\%
overhead and can be tuned depending on the users preference but, more
importantly, the API gives users fine-grain control of their
consistency/durability to support current practices or experimental workflows.}
